To be decided\\ Measuring Interaction before developing prototypes\\ Measuring Interaction Graphs

### 2015-09-10 paper.tex
Simon's initial Changes and comments to Abstract and Introduction.

% Load basic packages
\usepackage{balance}  % to better equalize the last page
\usepackage{graphics} % for EPS, load graphicx instead 
%\usepackage[T1]{fontenc}
\usepackage{txfonts}
\usepackage{times}    % comment if you want LaTeX's default font
\usepackage[pdftex]{hyperref}
% \usepackage{url}      % llt: nicely formatted URLs
\usepackage{color}
\usepackage{textcomp}
\usepackage{booktabs}
\usepackage{ccicons}
\usepackage{todonotes}

@@

\begin{abstract}
  Early development of prototypes and testing them with users is a good
  practice for user interfaces development. However, prototypes have
  to cover specific usage scenarios, and because they are limited
  in focus, the whole picture of the user interface is easily lost. Even
  simple questions dealing with numerosity and length of
  optimal execution paths or impact of possible user errors can be
  answered only for the specific scenarios being analyzed.

  We discuss a tool that transforms models of the behavior of a user
  interface into a graph. This is then used to specify usage
  scenarios, and to generate possible execution traces. Metrics based
  on number and length of different possible execution paths, with or
  without possible mistakes, can be easily computed. We apply the tool
  and metrics to well known examples of web mail apps, and show that
  important conclusions can be drawn even before the first prototype
  is built.

@@

\keywords{Authors' choice; of terms; separated; by semi\-colons;
  commas, within terms only; this section is required.}

\category{H.5.m.}{Information Interfaces and Presentation
  (e.g. HCI)}{Miscellaneous} \category{See
  \url{http://acm.org/about/class/1998/} for the full list of ACM
  classifiers. This section is required.}{}{}

\section{Introduction}

Developing good user interfaces for web or mobile applications is a
complex and expensive endeavor. One reason is the combination of
devices, interaction modalities and workflows that need to be
supported.

Adopting Usage Centered Development practices is effective, as is
following established design principles~\cite{constantine99}. Early
prototyping~\cite{buxton07} to explore part of the 5-dimensional
prototyping space~\cite{mccurdy06} is one of the most effective
techniques, especially when paired with usability investigations based
on user testing or heuristic evaluations.

@@

A designer, while conceiving and developing a solution, might benefit
from answers to seemingly simple questions that should not require a
significant investment of time and effort. Given one or more potential
solutions and some usage scenarios, interesting questions could
include: ``What are the optimal execution paths?'' or ``How many such
paths exist?'', or ``Are usage mistakes recoverable?''. At the moment,
even these straightforward questions are quite complex to answer. In
fact, they require inspection of prototypes, manual tracking of which
screens and widgets are used at which stage, and exhaustive searches.

@@

In this paper we present an approach that allows designer to quickly
be able to get these kinds of answers, and present it on a case study
comparing four different web mail front ends. We have chosen this
domain because it is very well understood by readers and yet those
trivial questions lead to non trivial results.

The approach is based on UML statechart models of the user interfaces
which are automatically processed to produce \emph{interaction
  graphs}. These are then used to unfold \emph{execution traces} that
are dependent on the specific usage scenarios being considered in the
analysis. On traces several graph-theoretic computations can be
performed to produce a dhasboard of different results that provide the
answers. Except for development of models and specification of the
desired scenarios, the other steps are totally automatic; models of a
design could be developed in a matter of a couple of hours.

>>>>

% Load basic packages
\usepackage{balance}  % to better equalize the last page
\usepackage{graphics} % for EPS, load graphicx instead 
%\usepackage[T1]{fontenc}
\usepackage{txfonts}
\usepackage{times}    % comment if you want LaTeX's default font
\usepackage[pdftex]{hyperref}
% \usepackage{url}      % llt: nicely formatted URLs
\usepackage{color}
\usepackage{textcomp}
\usepackage{booktabs}
\usepackage{ccicons}
\usepackage{todonotes}
\usepackage{soul}
\usepackage{paralist}

\usepackage{caption}
\usepackage{subcaption}
\usepackage[super]{nth}
\usepackage{amsmath}

@@

\begin{abstract}
  Early development and testing of prototypes is good
  practice for user interface development. However, prototypes have
  to cover specific usage scenarios, and because they are limited
  in focus, the whole picture of the user interface is easily lost. Even
  simple questions dealing with numerosity and length of
  optimal execution paths or impact of possible user errors can be
  answered only for the specific scenarios being analysed.

  We discuss a tool that transforms models of the behaviour of a user
  interface into a graph. This is then used to specify usage
  scenarios, and to generate possible execution traces. Metrics based
  on number and length of different possible execution paths, with or
  without possible mistakes, can be easily computed. We apply the tool
  and metrics to well known examples of web mail apps, \hl{and show that
  important conclusions can be drawn even before the first prototype
  is built. --- \textbf{replace with a real example statement}}.

@@

\keywords{Experimental; Evaluation; Statecharts: UML; UML-IDEA; Testing.}

\category{H.5.1}{Information interfaces and presentation (e.g., HCI)}{Multimedia Information Systems}. 
\category{H.5.2}{Information interfaces and presentation (e.g.,
  HCI)}{User Interfaces}.

\section{Introduction}
We present an approach that allows a designer to quickly 
\begin{inparaenum}[(a)]
\item understand how prone an application is to user navigation errors;  
\item understand how recoverable the application is from those errors;
\item support task analysis and task / scenario design;
\item support interaction design (IxD) in terms of consistency, error-proneness, and recovery premature commitment;
\end{inparaenum}
and further, allows different designs to be objectively compared to support designers in evaluating interaction sequences. To explain our work we present a case study
comparing four different web mail front ends. We have chosen this
domain because it is very well understood by readers and yet those
trivial questions lead to non trivial results.

The approach is based on UML state-chart models of the user interfaces
which are automatically processed to produce \emph{interaction graphs}. These are then used to unfold \emph{execution traces} that
are dependent on the specific usage scenarios being considered in the
analysis. On traces several graph-theoretic computations can be
performed to produce a dashboard of different results that provide the
answers. Except for development of models and specification of the
desired scenarios, the other steps are totally automatic; models of a
design could be developed in a matter of a couple of hours.

Developing good user interfaces for web or mobile applications is a
complex and expensive endeavour. One reason is the combination of
devices, interaction modalities and workflows that need to be
supported.

Adopting Usage Centred Development practices is effective, as is
following established design principles~\cite{constantine99}. Early
prototyping~\cite{buxton07} to explore part of the five-dimensional
prototyping space~\cite{mccurdy06} is one of the most effective
techniques, especially when paired with usability investigations based
on user testing or heuristic evaluations.

@@

A designer, while conceiving and developing a solution, might benefit
from answers to seemingly simple questions that should not require a
significant investment of time and effort. Given one or more potential
solutions and some usage scenarios, interesting questions could
include: \hl{``What are the optimal execution paths?'' or ``How many such
paths exist?'', or ``Are usage mistakes recoverable?'' -- \textbf{these seem to be software `executionist' and not so much user/designer is -- maybe we can say something more designer and save these to the technical bit of the paper?}}. At the moment,
even these straightforward questions are quite complex to answer. In
fact, they require inspection of prototypes, manual tracking of which
screens and widgets are used at which stage, and exhaustive searches.

@@
